Reconstructing Propositional Proofs in Type Theory

### 2017-10-27 paper/sections/type-theory.tex
[ paper ] type theory section additions.

Therefore, the task to verify a proof that proves a certain theorem is
translated to verify if the associated tree of the proof has the correspondent
type of the theorem. This process is known as \emph{type-checking}. This feature
of type theory allow us to verify a proof generated by \ATPs by reconstructing
its proof-term to type- check the proof. However, type theory has some
limitations for proof- reconstruction as we describe in the following section.

@@

We follow the same approach presented in~\cite{Altenkirch2015} to work with
propositional logic in type theory. Therefore, we include in the constructors
for theorems, structural rules and formation rules of all connectives in
Fig.~\ref{fig:CPL-inference-rules}.

@@

Since we pretend to reconstruct proofs generated by \Metis, and \Metis works
in classical logic,  we have included as an axiom, the
principle of excluded third (\abbre{PEM}).

Assuming \abbre{PEM}, we can justify refutation proofs.
We derive the \emph{reductio ad absurdum} rule (\abbre{RAA}) as the
formulation of the principle of proof by contradiction, that is, a derivation
of a contradiction, $⊥$, from the hypothesis $(¬ φ)$, is a
derivation of $φ$.

>>>>

Therefore, the task to verify a proof that proves a certain theorem is
translated to verify if the associated tree of the proof has the correspondent
type of the theorem. We know this process as \emph{type-checking}. This feature
of type theory allow us to verify a proof generated by \ATPs by reconstructing
its proof-term to type-check the proof. For such a purpose, we use a proof
assistant in type theory like \Agda to delegate this task.  However, this
approach has some weak points discussed in~\cite{Bezem2002}. Some limitations
from the type theory point of view for proof-reconstruction is described in the
following section.

@@

To reconstruct proofs generated by \Metis, we have formalized in type
theory the propositional logic (we briefly describe classical propositional logic in Appendix~\ref{app:CPL}).

To accomplish that formalization, the reader must take into account,
type theory and intuitionistic logic are based on the same basis of
the spirit of constructive mathematics. Therefore, proving propositions
demands a construction of the proof but in classical logic, they may
no exist and some proofs are proved by refutation. That is, using
a non-constructive method.

Therefore, to reconstruct \Metis derivations, we included as an axiom, the
principle of excluded third (\abbre{PEM}) since \Metis proves by refutation.
Assuming \abbre{PEM}, we can justify refutation proofs by deriving from it the
\emph{reductio ad absurdum} rule (\abbre{RAA}). The \abbrev{RAA} rule is the
formulation of the principle of proof by contradiction, that is, a derivation of
a contradiction, $⊥$, from the hypothesis $(¬ φ)$, is a derivation of $φ$.

@@

We implemented in~\cite{AgdaProp} the syntactical consequence relation in a
similar way as it was presented in~\cite{Altenkirch2015}. For that reason, we
have included structural rules like \emph{weaken} and formation rules for
connectives as the constructors for new theorems in
Fig.~\ref{fig:CPL-inference-rules}.
